---
title: "Bootstrap confidence intervals and diagnostics"
author: "Juwel Rana"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrap confidence intervals and diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
library(tvcQGComp)
```

## Bootstrap uncertainty

Use `tvcQGComp_survival_boot()` after finalizing the same `data`, `config`, and
intervention scenarios used for the point estimate. Resampling is performed at
the subject level to preserve each subject's longitudinal history.

```{r bootstrap}
if (FALSE) {
boot <- tvcQGComp_survival_boot(
  data = dat,
  config = cfg,
  n_boot = 500,
  mc_size = 10000,
  intervention_scenarios = scenarios,
  seed = 1234,
  parallel = TRUE,
  n_workers = 4,
  batch_size = 4,
  checkpoint_file = "bootstrap_checkpoint.rds",
  resume_from_checkpoint = TRUE,
  verbose = TRUE,
  stop_on_error = FALSE
)

boot_summary <- summarize_tvcqgcomp_bootstrap(boot)
boot_summary
}
```

Use a small number of replicates only to test the pipeline. The final analysis
should use enough successful replicates to estimate percentile limits with
adequate stability. Report the requested number, successful number, random
seed, Monte Carlo size, and interval method.

Checkpointing is particularly useful for long analyses. Store the checkpoint
outside the package source directory when building public releases.

## Run-level checks

```{r run-diagnostics}
if (FALSE) {
run_checks <- diagnose_tvcqgcomp_run(fit)
run_checks
}
```

Review warnings, missing outputs, scenario completeness, and numerical
irregularities before interpreting the effect estimates.

## Natural-course drift

When the analysis includes a simulated natural course, compare its
time-varying covariate trajectories with the observed data.

```{r drift}
if (FALSE) {
drift <- diagnose_tvcqgcomp_drift(
  observed_data = dat,
  run_obj = fit
)

drift$categorical_summary
drift$continuous_summary

plot_tvcqgcomp_drift(
  drift_obj = drift,
  continuous_stat = "mean",
  categorical_stat = "level_proportion"
)
}
```

Material drift can indicate model misspecification, unsupported simulated
histories, incorrect variable types, or unstable extrapolation. Diagnostics
should inform model revision; they are not a substitute for prespecifying the
causal assumptions and intervention definitions.

## Suggested reporting checklist

- Define the longitudinal time scale and event indicator.
- List all mixture components and quantization levels.
- Describe the natural-course mode and intervention scenarios.
- State each requested second-stage meta-model target.
- Report Monte Carlo size and bootstrap settings.
- Summarize failed bootstrap replicates and diagnostic findings.
