---
title: "Complete tvcQGComp analysis workflow"
author: "Juwel Rana"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Complete tvcQGComp analysis workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
render_outputs <- identical(
  tolower(Sys.getenv("TVCQGCOMP_RENDER_OUTPUTS", unset = "false")),
  "true"
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = render_outputs,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5
)
options(width = 90)
```

## Overview

`tvcQGComp` estimates the survival effect of joint interventions on a
quantized time-varying exposure mixture. The core workflow is:

1. arrange the observed data in long person-period format;
2. define the outcome, exposures, covariates, histories, and estimand;
3. validate the data against that configuration;
4. generate quantized intervention scenarios;
5. run Monte Carlo g-computation; and
6. interpret the point estimates and risk trajectories;
7. inspect run-level and natural-course diagnostics; and
8. use a subject-level bootstrap for confidence intervals.

This public release focuses on main-effect survival analyses with quantized
exposures.

## Required data structure

The input must contain one row per subject and follow-up interval. At minimum,
the configured columns must identify:

| Role | Example |
|---|---|
| Subject identifier | `UniqID` |
| Interval start and end | `TimeInn`, `TimeOut` |
| Interval event indicator | `status` |
| Time-varying exposures | `BC`, `NIT`, `SO4`, `NH4`, `OM` |
| Time-varying covariates | `CSize`, `Urban_form`, `CanadianRegion`, neighborhood indices |
| Baseline covariates | `sex`, education, employment, and other demographic variables |

Subject identifiers and interval times should uniquely identify rows. Base
variables used by model formulas must be present, correctly typed, and
supported at every time required by the simulation. Lagged formula terms do
not need to be stored in the input when `auto_history = TRUE`.

## Configure the analysis

The package includes the complete `toy_data` dataset used in this example.

```{r configure}
library(tvcQGComp)

data("toy_data", package = "tvcQGComp")
dat <- toy_data
dat$TimeOut <- as.integer(as.character(dat$TimeOut))

dim(dat)
head(dat[c("UniqID", "TimeInn", "TimeOut", "status", "BC", "NIT")])

exposures <- c("BC", "NIT", "SO4", "NH4", "OM")
active_tvc <- c(
  "Urban_form", "CanadianRegion", "CSize", "deprivation",
  "dependency", "instability", "ethnicconcentration"
)
baseline <- c(
  "age", "income_inadequacy", "sex", "visible_minority",
  "IndigenousIdentity", "marsth", "Education", "employment", "Occupation"
)

# These are formula term names only. auto_history creates the columns.
active_lag_terms <- paste0(active_tvc, "_lag1")
tvc_formulas <- setNames(lapply(active_tvc, function(response) {
  reformulate(
    c(
      "TimeOut", "I(TimeOut^2)", baseline,
      setdiff(active_lag_terms, paste0(response, "_lag1"))
    ),
    response = response
  )
}), active_tvc)

exposure_lag_terms <- paste0(exposures, "_lag1")
exposure_formulas <- setNames(lapply(exposures, function(response) {
  reformulate(
    c(
      "TimeOut", "I(TimeOut^2)", baseline, active_tvc,
      exposure_lag_terms
    ),
    response = response
  )
}), exposures)

cfg <- make_tvcqgcomp_config(
  id = "UniqID",
  time_in = "TimeInn",
  time_out = "TimeOut",
  outcome = "status",
  exposures = exposures,
  time_varying_covariates = active_tvc,
  time_fixed_covariates = baseline,
  factor_vars = c("income_inadequacy", active_tvc, baseline),
  outcome_formula = reformulate(
    c(exposures, "TimeOut", "I(TimeOut^2)", baseline, active_tvc),
    response = "status"
  ),
  tvc_formulas = tvc_formulas,
  exposure_formulas = exposure_formulas,
  exposure_types = setNames(rep("categorical", length(exposures)), exposures),
  covtypes = c(
    setNames(rep("categorical", length(active_tvc)), active_tvc)
  ),
  q = 4,
  exposome = "quantized",
  natural_course = c("calibration", "modeled_exposome"),
  auto_history = TRUE,
  baselags = TRUE,
  meta_target = c("HR", "RR", "RD")
)

data_valid <- validate_tvcqgcomp_data(dat, cfg)
data_valid
```

The configuration is the analysis contract. It should be finalized before
fitting the point estimate or bootstrap so both stages use identical models and
interventions. In this example, the seven neighborhood and urban-form
covariates and the five quantized exposures are modeled forward for the natural
course.

## Define interventions

The input exposures are continuous. The package determines their quartile
cutpoints internally, and the default helper assigns every internally
quantized exposure to each level in turn.

```{r interventions}
scenarios <- generate_intervention_scenarios(cfg)
names(scenarios)
```

Inspect `scenarios` before fitting. The scenario names become labels in the
result tables and plots.

## Fit the point estimate

The website executes this demonstration with a deliberately small Monte Carlo
sample so each code block can be followed by its actual output. Increase
`mc_size` substantially for the final analysis; the detailed implementation
uses 10,000.

```{r fit}
demo_mc_size <- 500L

fit <- tvcQGComp_survival(
  data = dat,
  config = cfg,
  mc_size = demo_mc_size,
  intervention_scenarios = scenarios,
  seed = 1234,
  verbose = FALSE
)

fit$risk_trajectory
fit$meta_effect_summary
```

The returned object contains fitted nuisance models, simulated intervention
data, risk summaries, natural-course outputs when requested, and the pooled
second-stage model.

`fit$meta_effect_summary` is the main point-estimate table, with one row for
each requested target. HR and RR estimates are reported on the ratio scale;
RD estimates are reported on the absolute risk-difference scale. Each estimate
represents a one-unit increase in the joint quantile intervention across all
mixture components. These are the non-bootstrap estimates; percentile
confidence intervals are added using the bootstrap later in this workflow.

## Inspect risk trajectories

```{r curves}
risk_curves <- as_risk_curve_df(fit)
head(risk_curves)

plot_cumulative_risk_trajectory(fit)
```

Risk trajectories are useful for understanding when scenarios begin to
separate. Final HR, RR, and RD estimates come from their configured
second-stage meta-models.

## Check the fitted run

Before interpreting the effect estimates, check the structure and numerical
behavior of the fitted run.

```{r run-diagnostics}
run_checks <- diagnose_tvcqgcomp_run(fit)
run_checks$summary
run_checks$scenario_exposure_checks
run_checks$final_risk_table
```

Review warnings, missing outputs, scenario completeness, Monte Carlo panel
balance, exposure constancy under static interventions, and numerical
irregularities. Resolve substantive problems before proceeding to inference.

## Check natural-course drift

The natural-course simulation should reproduce the observed trajectories of
the modeled time-varying covariates reasonably well. Because the configuration
requested `modeled_exposome`, `fit$natural` contains forward-simulated
time-varying covariates and exposures. Supply the observed data as the
comparison target; the fitted object supplies the natural-course simulation
and configuration.

```{r drift-diagnostics}
drift <- diagnose_tvcqgcomp_drift(
  observed_data = dat,
  run_obj = fit
)

drift$categorical_summary
drift$continuous_summary

plot_tvcqgcomp_drift(
  drift_obj = drift,
  continuous_stat = "mean",
  categorical_stat = "level_proportion",
  main = "Observed versus simulated natural course"
)
```

For categorical variables, the summary reports differences in time-specific
level proportions. For continuous variables, it reports differences in means,
standard deviations, and selected quantiles. Material drift can indicate model
misspecification, unsupported simulated histories, incorrect variable types,
or unstable extrapolation. These diagnostics should guide model revision; they
do not replace the causal assumptions required for identification.

## Bootstrap uncertainty

Use the same data, configuration, and intervention scenarios as the point
estimate. Resampling is performed at the subject level to preserve each
participant's longitudinal history.

The website uses five replicates only to demonstrate the returned output. This
is not sufficient for inference; use a substantially larger number, such as
500, for the final analysis.

```{r bootstrap}
demo_n_boot <- 5L

boot <- tvcQGComp_survival_boot(
  data = dat,
  config = cfg,
  n_boot = demo_n_boot,
  mc_size = demo_mc_size,
  intervention_scenarios = scenarios,
  seed = 1234,
  parallel = FALSE,
  checkpoint_file = NULL,
  verbose = FALSE,
  stop_on_error = FALSE
)

boot_summary <- summarize_tvcqgcomp_bootstrap(boot)
boot_summary
```

Use a small number of replicates only when testing the pipeline. The final
analysis should use enough successful replicates to estimate percentile limits
with adequate stability. Store checkpoint files outside the package source
directory when preparing a public release.

## Suggested reporting checklist

- Define the longitudinal time scale and event indicator.
- List all mixture components and quantization levels.
- Describe the natural-course mode and intervention scenarios.
- State each requested second-stage meta-model target.
- Report the Monte Carlo size, bootstrap settings, and random seeds.
- Report the requested and successful numbers of bootstrap replicates.
- Summarize run-level checks, natural-course drift, and any model revisions.
